---
title: "R_Programming_week_2_IP_(CLUSTERING)"
author: "Ruth Nguli"
date: '2022-03-24'
output: pdf_document
---
## Defining the Question

To understand customer’s behavior from data that collected over the past year. 

## Metric for success

The metric of success will be attained on identifying the characteristics of customer groups.

## Understanding the business context

Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia.

## Experimental Design

Define the question, the metric for success, the context, experimental design taken.

Read and explore the given dataset.

Cleaning Data

Perform Exploratory Data Cleaning (Univariate & Bivariate)

create a Clustering model

Conclusion

Recommendations

## Data Source

The dataset source link: http://bit.ly/EcommerceCustomersDataset

## Reading Data

```{r}
# Loading and reading data from source link
#
library(tidyverse)
brand <- read.csv("http://bit.ly/EcommerceCustomersDataset")

# Previewing the head of data
#
head(brand)
```
```{r}
# Previewing tail of data
#
tail(brand)

```
```{r}
# Checking the data structure 
#
str(brand)
```
*The dataset has 12330 observations (rows) and 18 variables(columns)
*The data types are appropriately represented

```{r}
# Looking at the columns
#
library(tidyverse)
colnames(brand)
```
## Data Cleaning

```{r}
# looking at data summary 
#
summary(brand)

```
* there are some NA values which will need be imputed

```{r}
## Calling Amelia and mice libraries for data imputation
#
library(Amelia)
library(mice)
library(VIM)
```
```{r}
## Confirming if the data has missing
#
mapply(anyNA,brand)
```

```{r}
# Imputing the missing values by predicting missing values with mice package
# 

miss_mod <- mice(brand[, c("Administrative" , "Administrative_Duration", "Informational", "Informational_Duration", "ProductRelated", "ProductRelated_Duration", "BounceRates", "ExitRates")], method='rf')

completed <- complete(miss_mod)

```
```{r}
# placing predicted missing values into the main data set
# ---
# 
brand$Administrative <- completed$Administrative
brand$Administrative_Duration <- completed$Administrative_Duration
brand$Informational <- completed$Informational
brand$Informational_Duration <- completed$Informational_Duration
brand$ProductRelated <- completed$ProductRelated
brand$ProductRelated_Duration <- completed$ProductRelated_Duration
brand$BounceRates <- completed$BounceRates
brand$ExitRates <- completed$ExitRates

```

```{r}
# confirming if there no more missing values
# 
anyNA(brand)
```
```{r}
# checking duplicates
#
anyDuplicated(brand)

```
*presence of duplicates

```{r}
# Dropping duplicated
#
brand <- brand %>% distinct()

# confirming if duplicates were successfully dropped
#
anyDuplicated(brand)
```
*Successfully dropped duplicates in the data*

```{r}
# checking outliers
#
non_cat <- brand %>% select("Administrative" , "Administrative_Duration", "Informational", "Informational_Duration", "ProductRelated", "ProductRelated_Duration", "BounceRates", "ExitRates", "PageValues", "SpecialDay", "OperatingSystems",  "Browser", "Region", "TrafficType")
boxplot(non_cat)

```
*There are noticeable outliers which represent real time data.

## Univariate Analysis

```{r}
# Plotting a Histograms of Administrative and duration
#
attach(brand)
hist(Administrative, col="pink")
hist(Administrative_Duration, col="pink")
```

*Administrative is skewed to the right. Majority of the visitors did not visit the Administrative page.
*Majority of the visitors spend less time on the page.


```{r}
# Plotting a Histograms of Informational and duration
#
hist(Informational, col="pink")
hist(Informational_Duration, col="pink")

```

* Informational and Informational_Duration are skewed to the right.
* most visitors did not visit the Informational page. 
* most visitors who visited the page spend less time on the page.

```{r}
## Plotting  Histograms of Product related and duration
#
hist(ProductRelated, col="pink")
hist(ProductRelated_Duration, col="pink")
```

* ProductRelated and ProductRelated_Duration are skewed to the right.
* most visitors did not visit the ProductRelated page. 
* most visitors who visited the page spend less time on the page.

```{r}
## Plotting  Histograms of BounceRate and ExitRate
#
hist(BounceRates, col="pink")
hist(ExitRates, col="pink")
```
* Bounce rate at 0.00.has the highest frequency.
* ExitRate was happening between 0.00 and 0.05 rate.

```{r}
## Plotting  Histograms of Pagevalue and Special day
#
#
ggplot(brand, aes(x=reorder(SpecialDay, SpecialDay, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='SpecialDay')
```

```{r}
## Plotting  Bar plot of Month
#
ggplot(brand, aes(x=reorder(Month, Month, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Month')

```
* most customers visited the pages in the months of May,Nov,March and Dec.

```{r}
## Plotting  Bar plot of OperatingSystems
#
ggplot(brand, aes(x=reorder(OperatingSystems, OperatingSystems, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Operating Systems')
```
* most visitors used 2 different operating systems to access the brand's pages.

```{r}
## Plotting  Bar plot of Browser
#
ggplot(brand, aes(x=reorder(Browser, Browser, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='browser')
```
* most visitors used 2 different Browsers to access the brand's pages.

```{r}
## Plotting  Bar plot of Region
#
ggplot(brand, aes(x=reorder(Region, Region, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Region')
```
* most customers accessed the brand's pages from the one region.

```{r}
## Plotting  Bar plot of TrafficType
#
ggplot(brand, aes(x=reorder(TrafficType, TrafficType, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='TrafficType')

```
* Most of the visitors had between 1 & 3 traffic types.

```{r}
## Plotting  Bar plot of VisitorType
#
ggplot(brand, aes(x=reorder(VisitorType, VisitorType, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Visitor Type')
```
* Majority of page visitors were returning visitors

```{r}
## Plotting  Bar plot of Weekend
#
ggplot(brand, aes(x=reorder(Weekend, Weekend, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Weekend')
```
* we observe that most page visits were on weekdays.

```{r}
## Plotting  Bar plot of Revenue
#
ggplot(brand, aes(x=reorder(Revenue, Revenue, function(x)-length(x)))) +
geom_bar(fill='pink') +  labs(x='Revenue')
```
* We observe most of the visits in the brand's pages did not generate any revenue.

## Bivariate Analysis

```{r}
# Administrative Vs Revenue
#
ggplot(brand,aes(Administrative, Revenue, colour= Revenue))+
  geom_step(size=3)

```

```{r}
# Informational Vs Revenue
#
ggplot(brand,aes(Informational, Revenue, colour= Revenue))+
  geom_step(size=3)

```
```{r}
# ProductRelated Vs Revenue
#
ggplot(brand,aes(ProductRelated, Revenue, colour= Revenue))+
  geom_step(size=3)

```
* We observe that most of Administrative,Informational & ProductRelated page visits did not generate much revenue.

```{r}
# ExitRate Vs Revenue
#
ggplot(brand,aes(ExitRates, Revenue, colour= Revenue))+
  geom_point(size=3)

```
```{r}
# Bouncerate Vs Revenue
#
ggplot(brand,aes(BounceRates, Revenue, colour= Revenue))+
  geom_point(size=3)

```
```{r}
# Pagevalues Vs Month
#
ggplot(brand,aes(PageValues, Month, colour= Revenue))+
  geom_point(size=3)

``` 
* Page value mostly attracted revenue through out the year.

```{r}
# Month Vs Specialday,  Revenue
#
ggplot(brand,aes(SpecialDay, Month, colour= Revenue))+
  geom_point(size=3)

```
* Revenue was generated on special day in the months of June and July

```{r}
# OperatingSytems Vs Month
#
ggplot(brand,aes(OperatingSystems, Month, colour= Weekend))+
  geom_point(size=3)

``` 
* 
```{r}
# Pagevalues Vs Month
#
ggplot(brand,aes(VisitorType, Month, colour= Revenue))+
  geom_point(size=3)

``` 
## Modelling

```{r}
# converting months to numeric
#
brand$Month <- match(Month,month.abb)
tail(brand)
```

```{r}
# Removing class label before label encoding
# 
df <- brand[,c(1:17)]
head(df)


```



```{r}
# Standardising the data 
#
data <- df %>% select(-VisitorType, -Weekend, -Month) %>%  scale()
head(data)
```
## K-Means Clustering

```{r}
# Determining the optimal value of k
#
library(factoextra)
fviz_nbclust(data, kmeans, method= "silhouette")
```
* The optimal k value is 3.

```{r}
# Applying the K-means at k=3
# ---
# 
k_clust<- kmeans(data,3) 

# Previewing the no. of records in each cluster
# 
k_clust$size


```
```{r}
# Visualizing the  clustering results
#
fviz_cluster(kmeans(data, centers = 3), data = data)
```
```{r}
# displaying the results in a table
#
brand.class<- brand[, "VisitorType"]
table(k_clust$cluster, brand.class)
```

## Hierarchical

```{r}
# computing the Euclidean distance between observations, 
#
df <- dist(data, method = "euclidean")

# hierarchical clustering using the complete method
# 
hc <- hclust(df, method = "complete" )

hc

```

```{r}
# plotting adendrogram
# 
plot(hc)
rect.hclust(hc, k = 10, border = 2:5)

```

## DBSCAN Clustering

```{r}
# getting optimum eps
#
library(fpc)
library(dbscan)
kNNdistplot(data, k=3) 
abline(h = 3, lty=2)
```

```{r}
# Applying dbscan algorithm with the optimal eps = 3
#
db <- dbscan(data,eps=3,MinPts = 4)
db
```

```{r}
# Plotting DBSCAN
#

hullplot(data,db$cluster)
```

```{r}
#dispalying the cluster results in a table

table(db$cluster, brand.class)
```
## conclusion

We conclude that returning visitors are most likely to generate revenue 
Pagevalue through out the year showed revenue was being genenaated.

