---
title: "R-Programming_EDA_Week1_IP"
author: "RuthNguli"
date: '2022-03-18'
output:
  html_document:
    df_print: paged
---


## Defining the Question

To identify which individuals are most likely to click on an online cryptography course ads.

## Metric for success

The metric of success will be attained on identifying individuals who click on the ads.

## Understanding the business context

Cryptography is an indispensable tool for protecting information in computer systems. A cryptography course teaches how the cryptographic system a works and its real world application.  

## Experimental Design

Define the question, the metric for success, the context, experimental design taken.

Read and explore the given dataset.

Cleaning Data

Perform Exploratory Data Cleaning (Univariate & Bivariate)

Modelling

Conclusion

Recommendations

## 1. Reading data
```{r}
# read data from url: http://bit.ly/IPAdvertisingData
# load data
#
library(tidyverse)
ads <- read.csv("http://bit.ly/IPAdvertisingData")

# preview head of the data
#
head(ads)
```
```{r}
# previewing the tail of the data
#
tail(ads)

```
```{r}
# checking column names
#
colnames(ads)
```
```{r}
# Checking the data has appropriate data types
#
str(ads)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

*The advertising data has 1000 rows and 10 columns*
*Column Male and Clicked.on.Ad are represented as integer however should be converter to factor as they are represnting categorical variables*


## Data Cleaning
```{r}
# change data types of Male and Clicked.on.Ad columns from int to factor
#
ads$Male <- as.factor(ads$Male)
ads$Clicked.on.Ad <- as.factor(ads$Clicked.on.Ad)

# confirming if the changes have made successfully
#
str(ads)
```
```{r}
#Extracting and creating Date column from timestamp
#
ads$Date <- as.Date(ads$Timestamp)
glimpse(ads)

```


```{r}
# Checking missing values
#
library(Amelia)
missmap(ads)
```
*Our dataset does not have any missing values*

```{r}
# Checking for any duplicates
#
sum(duplicated(ads))
```
*There are no duplicates*

```{r}
# Checking for outliers
# 
non_char <- ads %>% select(Daily.Time.Spent.on.Site, Age, Area.Income, Daily.Internet.Usage, Male,Clicked.on.Ad, Date)
boxplot(non_char)
```
*There a few identifiable outliers in Area.Income column, we leave them as they represent real data.*
```{r}
# Checking the summarry of dataset
#
summary(ads)
```
## Bivariate analysis
```{r}
# Attaching ads data to R 
#
attach(ads)
# Plotting a histogram of Daily Time spent on site
hist(Daily.Time.Spent.on.Site, col='pink')
```
*Most time spent on site is between 65 and 80*
```{r}
#Plotting a histogram of Age
#
hist(Age, col='pink')
```
*Most participants are aged between 25 and 40 years old*
```{r}
# plotting a histogramof Area income
#
hist(Area.Income, col="pink")
```
*Income is skewed to the left. Most participants had income between 50,000 and 70,000*

```{r}
# A bar plot of Male Partipation
#
barplot(table(Male), col="pink", main="Bar Plot of Male distribution")
```
*Male participants were slightly fewer than those not male*
```{r}
# A bar plot of Clicked on Ads
#
barplot(table( Clicked.on.Ad), col="pink", main="A Bar plot of Clicked on Ads")
```
*There is equal distribution between those who clicked and those didn't click on AD*

```{r}
# Plotting a Histogram of Daily internet usage
#
hist(Daily.Internet.Usage, col="pink")
```
## Bivariate Analysis
```{r}
colnames(ads)
```

```{r}
# A Boxplot of clicked on ad vs Time spent on ad
#
plot(Daily.Time.Spent.on.Site ~ Clicked.on.Ad, data = ads, col="pink", main="A Box Plot of Daily time spent on site vs Clicked on Ad")
```
* Most people who clicked on ads did not spent much time on site*
```{r}
# A scatter-plot of Age Vs Clicked on ads
#
ggplot(ads,aes(Age,Clicked.on.Ad, colour= Clicked.on.Ad))+
  geom_point(size=3)
```
*Those who clicked on ads are aged between 20  and 60.*
```{r}
# A Scatter plot of Date Vs Daily time spent on site
#
ads %>% ggplot(aes(Date,Daily.Time.Spent.on.Site, colour = Clicked.on.Ad))+
  geom_point(size=3, alpha = 0.5)+geom_smooth(method=lm, se= F)
```
*Majority of those who clicked on ad through Jan to July spent less time online*
```{r}
# A scatter plot of Age VS time spent on site
#
ads %>% ggplot(aes(Age,Daily.Time.Spent.on.Site, colour = Clicked.on.Ad))+
  geom_point(size=3, alpha = 0.5)+geom_smooth(method=lm, se= F)
```
*Most people who clicked on ads are aged between 30 -50 years and they spent much less time online*
```{r}
# A scatter plot of random selected countries VS Ages
#
filter(ads, Country %in% c("Cuba","Tunisia","Korea","Peru","Thailand","Greece","Senrgal","Ukraine","Australia")) %>%
  ggplot(aes(Age,Country , colour=Clicked.on.Ad)) +
  geom_point()
```

*Most Countries the people who click on ads are 30yrs and above*
```{r}
# A scatter plot of Age VS time spent on site
#
ads %>% ggplot(aes(Area.Income,Daily.Time.Spent.on.Site, colour = Clicked.on.Ad))+
  geom_point(size=3, alpha = 0.5)+geom_smooth(method=lm, se= F)
```

```{r}
Num <- ads %>% select(Daily.Time.Spent.on.Site, Age, Area.Income, Daily.Internet.Usage)

corr <- cor(Num)

corr
```
*Daily internet usage is positively corr to daily time spent online 0.52 , Age is negatively corr to time spent online -0.37*

## Modelling
```{r}
ads$Clicked.on.Ad <- as.numeric(Clicked.on.Ad) 
```

```{r}
## KNN model
# performing normalization
normal <- function(x) (
  return( ((x - min(x)) /(max(x)-min(x))) )
)

data <- ads[1:10]

ads_norm <- as.data.frame(lapply(data%>% select(-Timestamp,-Country,-City,-Ad.Topic.Line,-Male,-Clicked.on.Ad), normal))
ads_norm
```
```{r}
set.seed(55) # to get same random sample

# selecting 70% of the data
ads_samp <- sample(1:nrow(ads_norm), size = nrow(ads_norm)*0.7,replace=FALSE)
```
```{r}
# getting 70% train and 30 % test data X(Independent)
X_train<- data[ads_samp,]
X_test <- data[-ads_samp,]

```


```{r}
# Creating train and test dataset for y (dependent)
y_train <- data[ads_samp,10]
y_test <- data[-ads_samp,10]

```


```{r}
library(class)
# Applying k-NN classification algorithm.
# No. of neighbours are generally square root of total number of instances
neigh<- round(sqrt(NROW(y_train)))+1 # here we want to have the number y_training data 
# Applying the knn algorithm
model<- knn(train = X_train,  test = X_test, cl = y_train, k = neigh) 


```

```{r}
# getting a confussion matrix
library(caret)

confusionMatrix(table(y_test, model))
```

```{r}
# Calculating the Accuracy
mean(y_test== model)*100
```

```{r}
# Tuning the model
i = 1  #initiating a loop
k.optm = 1
for (i in 1:25) {
  model<- knn(train = X_train,  test = X_test, cl=y_train, k=i)
  k.optm[i] <- mean(y_test== model)*100
  k=i
  cat(k ,"=", k.optm[i],'\n')   # to print accuracy
  
}

plot(k.optm, type="b", xlab="k- values", ylab="Accuracy levels") #to print optimum accuracy level
```


## Conclusion
Ads are mostly clicked: Those aged between 30 and 60 years
                        Those who spent less time online
                        Ads are clicked  through out the months.
                        
## Recommendations
Its Recommend that to target age group between 30 and 60, have a well detailed and well explained advert as those who click on ads do spent much time on site and make sure advert run through out the year